---
title: "Successful completion of residential substance treatment"
author: "Abby Stevens"
date: "2020-11-25"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)

library(tidyverse)
library(tidymodels)
library(reshape2)
library(broom)
library(pROC)
library(vip)
source('code/teds_utils.R')
```


In this experiment, we are interested in the factors that determine whether or not a person completes residential treatment for 5 substances of interest: Opioids, Alcohol, Cocaine, Marijuana, and Methamphetamine. For each substance, we tune a random forest model and compute the test error/variable importances.

```{r read discharge dataset}
mydata <- read.csv("data/tedsd_puf_2017.csv")
medexp <- read.csv("data/medicaid_expansion.csv")
# merge
mydata <- merge(mydata, medexp, by='STFIPS')

# filter by residential treatment
mydata <- mydata %>% filter(SERVICES %in% c(4,5))

# create "substance" flag
mydata$substance = 'Other'
mydata$substance = ifelse(mydata$SUB1 %in% c(5,6,7), 'Opioids', mydata$substance)
mydata$substance = ifelse(mydata$SUB1 == 2, 'Alcohol', mydata$substance)
mydata$substance = ifelse(mydata$SUB1 == 3, 'Cocaine', mydata$substance)
mydata$substance = ifelse(mydata$SUB1 == 4, 'Marijuana', mydata$substance)
mydata$substance = ifelse(mydata$SUB1 == 10, 'Methamphetamine', mydata$substance)

# define variables
myvars <- c("ROUTE1", "ALCFLG", "FREQ1", "FRSTUSE1", "IDU","COKEFLG", "BENZFLG", "PSYPROB", "HLTHINS", "PRIMPAY", "AGE", "GENDER", "LIVARAG", "DIVISION", "METHUSE", "NOPRIOR", "EDUC", "HERFLG","RACE","MARSTAT","PRIMINC","ETHNIC","ARRESTS_D", "MEDEXP", "SUB2")

# create response
mydata$COMPLETED = ifelse(mydata$REASON==1, 1, 0) #create response variable
mydata[myvars] <- as.data.frame(lapply(mydata[myvars], factor))

response = "COMPLETED"
```

We can observe the overall completion rates for the different substances:

```{r}
sbs_response_plots(mydata, c("substance"), response)
```

Now, we loop through the substances, train the model, and keep track of the results.


```{r}
set.seed(123) #for replicability
test_aucs = list()

for (sub in c("Alcohol", "Cocaine", "Marijuana", "Methamphetamine", "Opioids")){
  print(sub)
  teds0 <- mydata %>% filter(substance==sub)
  teds = teds0[, myvars]
  teds[, response] = teds0[, response]
  
  # split
  teds_split = initial_split(teds, prop=3/4)
  # extract training and testing sets
  teds_train <- training(teds_split)
  teds_test <- testing(teds_split)
  
  # fit logistic
  fm <- as.formula(paste(response, "~ ."))
  lg = glm(fm, family=binomial, data=teds_train)
  summary(lg)

test_prob = predict(lg, newdata = teds_test, type = "response")
test_roc = roc(teds_test$COMPLETED ~ test_prob, plot = TRUE, print.auc = TRUE)
  rf <- teds_rf(teds, myvars, response)
  test_aucs[sub] = rf$test_performance$.estimate[2]
  
  # variable importances
  teds[,response] = as.factor(teds[[response]])
  final_model <- fit(rf$model, teds)
  t = final_model %>% 
        pull_workflow_fit() %>% 
        vi()
  colnames(t) = c("Variable", sub)
     
  if(sub != 'Alcohol'){
     vidf = merge(vidf, t, by='Variable')
  } else{
    vidf = t
  }
}
```

First we take a look at the test AUCs for the different substances and see that for all substnaces it is between .75 and .8.
```{r}
test_aucs
```


Finally, we can compare variable importance across substances to see how they differ.

```{r}
ggplot(melt(vidf), aes(x = variable, y = Variable, fill = value)) +
  geom_tile()
```


